---
title: "Sample Genome Stats"
output:
  html_document:
    toc: true
    toc_float: true
    number_section: true
author: Mark Wadsworth
date: '2022-05-12'
params:
  runDRFDir: "."
  calculateBamStats: "./Calculate_Bam_Stats"
  genesOfInterest: "./genesOfInterest.bed"
  OutDir: "."
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE,
    collapse = TRUE
)
knitr::opts_knit$set(root.dir = params$OutDir)

require("BiocManager")

requiredPackages = c('gridExtra', 'reshape2','ggplot2','data.table', "org.Hs.eg.db", "GenVisR", "TxDb.Hsapiens.UCSC.hg38.knownGene", "BSgenome.Hsapiens.UCSC.hg38")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) BiocManager::install(p)
  library(p,character.only = TRUE)
}

print(getwd())
```

# Load RUN_DRF files

```{r}
AvgDepth.df = NULL
CumulativeBases.df = NULL
fileList = Sys.glob(paste0(params$runDRFDir, "/*GlobalStats.txt"))
#print(fileList)
for (file in fileList){
  #print(file)
  temp = read.table(file, sep="\t", header=T, row.names=1, stringsAsFactors = F)
  #print(temp[,"AvgDepth", drop=F])
  if (!is.null(AvgDepth.df)){
    newCols = c(colnames(AvgDepth.df), unlist(strsplit(basename(file), "_"))[1])
    AvgDepth.df = cbind(AvgDepth.df, temp[, "AvgDepth", drop=F])
    colnames(AvgDepth.df) = newCols
    
    CumulativeBases.df = cbind(CumulativeBases.df, temp[, "NumBasePairs", drop=F])
    colnames(CumulativeBases.df) = newCols
  } else{
    AvgDepth.df = as.data.frame(temp[, "AvgDepth", drop=F])
    colnames(AvgDepth.df) = unlist(strsplit(basename(file), "_"))[1]
    
    CumulativeBases.df = as.data.frame(temp[, "NumBasePairs", drop=F])
    colnames(CumulativeBases.df) = unlist(strsplit(basename(file), "_"))[1]
  }
}


```

# Sample Average Depth

## Average Depth per Chromosome

The average depth is calculated by calculating the sum from bedtools merge on the base-by-base coverage file from the 01-RUN_DRF step. This is passed to a python script that takes the sum of the per base coverage and divides it by the number of bases covered per chromosome or per genome. 

Relevant lines of code:

 * bin/calc_perBaseStats.py
 * modules/01-RUN_DRF: 154-181


```{r}
AvgDepth.df.melt = melt(as.matrix(AvgDepth.df))
AvgDepth.df.melt$Var1 = factor(AvgDepth.df.melt$Var1, levels = c("AllChrs", "chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22", "chrX", "chrY", "chrM"))


#ggplot(AvgDepth.df.melt[which(AvgDepth.df.melt$Var1 != "chrM"),], aes(x=Var1, y=value, fill=Var2)) + geom_bar(stat="identity", position="dodge")
#ggplot(AvgDepth.df.melt[which(AvgDepth.df.melt$Var1 != "chrM"),], aes(x=Var1, y=value, color=Var2)) + geom_jitter(width=.1)
ggplot(AvgDepth.df.melt[which(!AvgDepth.df.melt$Var1 %in% c("AllChrs", "chrM")),], aes(x=Var1, y=value)) + geom_boxplot() + geom_jitter(width=.1, aes(color=Var2)) + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.title= element_blank()) + ylab("Average Depth Per Chromosome") + xlab("") 
ggplot(AvgDepth.df.melt[which(!AvgDepth.df.melt$Var1 %in% c("AllChrs", "chrM", "chrX", "chrY")),], aes(x=Var1, y=value)) + geom_boxplot() + geom_jitter(width=.1, aes(color=Var2)) + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.title= element_blank()) + ylab("Average Depth Per Chromosome") + xlab("") 
#ggplot(AvgDepth.df.melt[which(AvgDepth.df.melt$Var1 != "chrM"),], aes(x=Var1, y=value)) + geom_boxplot() + geom_jitter(width=.1) + coord_flip()

```

## Average Depth per Sample

I don't have the median computed for the full genome because of time limitations.

```{r}
ggplot(AvgDepth.df.melt[which(AvgDepth.df.melt$Var1 == "AllChrs"),], aes(x=Var1, y=value)) + geom_boxplot() + geom_jitter(width=.1, aes(color=Var2)) + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("") + ylab("Average Depth Across All Chromosomes")

hist(AvgDepth.df.melt[which(AvgDepth.df.melt$Var1 == "AllChrs"),"value"], main="Average Depth per Sample", xlab = "Average Depth per Sample", ylab = "Number of Samples") 
```

## Number of Bases in Genome

This is the total number of bases in the version of the Autosomal genome used in this analysis. If there are more than one and the Standard Deviation is not 0 there is a problem.

```{r}
print(unique(unlist(CumulativeBases.df["AllChrs",])))
print(sd(CumulativeBases.df["AllChrs",]))
```

# Extraction Region Depth

## Load Extraction Region Data

```{r}
ExtractionDarkByDepthFiles = Sys.glob(paste0(params$calculateBamStats, "/*_low_depth_perRegionMetrics.bed"))
ExtractionDarkByMapQFiles = Sys.glob(paste0(params$calculateBamStats, "/*_low_mapq_perRegionMetrics.bed"))

ExtractionDarkByDepth.Avg.dt = NULL
ExtractionDarkByMapQ.Avg.dt = NULL
ExtractionDarkByDepth.Median.dt = NULL
ExtractionDarkByMapQ.Median.dt = NULL

for( file in ExtractionDarkByDepthFiles ){
  temp = fread(file, sep = "\t", header = F)
  temp$rn = paste0(temp$V1,":",temp$V2,"-",temp$V3)
  setkey(temp, rn)
  if(!is.null(ExtractionDarkByDepth.Avg.dt)){
    newColName = c(colnames(ExtractionDarkByDepth.Avg.dt), unlist(strsplit(basename(file), "_"))[1])
    ExtractionDarkByDepth.Avg.dt = merge(ExtractionDarkByDepth.Avg.dt, temp[,c("rn", "V5")], by="rn", all=T)
    colnames(ExtractionDarkByDepth.Avg.dt) = newColName
    
    ExtractionDarkByDepth.Median.dt = merge(ExtractionDarkByDepth.Median.dt, temp[,c("rn", "V6")], by="rn", all=T)
    colnames(ExtractionDarkByDepth.Median.dt) = newColName
    
  } else{
    ExtractionDarkByDepth.Avg.dt = temp[,c("rn","V4", "V5")]
    colnames(ExtractionDarkByDepth.Avg.dt) = c("rn", "ExtractionRegion", unlist(strsplit(basename(file), "_"))[1])
    
    ExtractionDarkByDepth.Median.dt = temp[,c("rn","V4", "V6")]
    colnames(ExtractionDarkByDepth.Median.dt) = c("rn", "ExtractionRegion", unlist(strsplit(basename(file), "_"))[1])
  }
  
}

for( file in ExtractionDarkByMapQFiles ){
  temp = fread(file, sep = "\t", header = F)
  temp$rn = paste0(temp$V1,":",temp$V2,"-",temp$V3)
  setkey(temp, rn)
  if(!is.null(ExtractionDarkByMapQ.Avg.dt)){
    newColName = c(colnames(ExtractionDarkByMapQ.Avg.dt), unlist(strsplit(basename(file), "_"))[1])
    ExtractionDarkByMapQ.Avg.dt = merge(ExtractionDarkByMapQ.Avg.dt, temp[,c("rn", "V5")], by="rn", all=T)
    colnames(ExtractionDarkByMapQ.Avg.dt) = newColName
    
    ExtractionDarkByMapQ.Median.dt = merge(ExtractionDarkByMapQ.Median.dt, temp[,c("rn", "V6")], by="rn", all=T)
    colnames(ExtractionDarkByMapQ.Median.dt) = newColName
  } else{
    ExtractionDarkByMapQ.Avg.dt = temp[,c("rn","V4", "V5")]
    colnames(ExtractionDarkByMapQ.Avg.dt) = c("rn", "ExtractionRegion", unlist(strsplit(basename(file), "_"))[1])
    
    ExtractionDarkByMapQ.Median.dt = temp[,c("rn","V4", "V6")]
    colnames(ExtractionDarkByMapQ.Median.dt) = c("rn", "ExtractionRegion", unlist(strsplit(basename(file), "_"))[1])
  }
  
}

ExtractionDarkByDepth.Avg.df = data.frame(ExtractionDarkByDepth.Avg.dt, row.names="rn")
ExtractionDarkByMapQ.Avg.df = data.frame(ExtractionDarkByMapQ.Avg.dt, row.names="rn")
ExtractionDarkByDepth.Median.df = data.frame(ExtractionDarkByDepth.Median.dt, row.names="rn")
ExtractionDarkByMapQ.Median.df = data.frame(ExtractionDarkByMapQ.Median.dt, row.names="rn")

```

## Dark-By-Depth

This is the intersection of Dark-by-Depth regions within Extraction regions.

Relevant Code:

 * modules/02-CALCULATE_BAM_STATS.nf : 124-127
 

```{r}
extractionDarkByDepth.Avg.melt = melt(as.matrix(ExtractionDarkByDepth.Avg.df[,2:ncol(ExtractionDarkByDepth.Avg.df)]))
extractionDarkByDepth.Median.melt = melt(as.matrix(ExtractionDarkByDepth.Median.df[,2:ncol(ExtractionDarkByDepth.Median.df)]))

ggplot(extractionDarkByDepth.Avg.melt[which(extractionDarkByDepth.Avg.melt$value != "."),], aes(x=as.numeric(as.character(value)), color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Average Depth") + ggtitle("Average Depth of Dark-By-Depth Extraction Regions")

#ggplot(extractionDarkByDepth.Median.melt[which(extractionDarkByDepth.Median.melt$value != "."),], aes(x=as.numeric(as.character(value)), color=Var2)) + geom_density() + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Median Depth") + ggtitle("Median Depth of Dark-By-Depth Extraction Regions")
```

## Dark-By-MapQ

These are the log10(average depth) and log10(median depth) of extraction the regions. These Dark-by-MapQ regions constitute regions in which variants can't be called because of multiple alignments. The rescue step rescues variants in these regions. If these have high depth you can have confidence that the run is sufficient to use in downstream analyses. This is the intersection of Dark-by-MapQ regions and the extraction CDS regions.

Relevant Code:

 * modules/02-CALCULATE_BAM_STATS.nf : 119-122

```{r}
extractionDarkByMapQ.Avg.melt = melt(as.matrix(ExtractionDarkByMapQ.Avg.df[,2:ncol(ExtractionDarkByMapQ.Avg.df)]))
extractionDarkByMapQ.Median.melt = melt(as.matrix(ExtractionDarkByMapQ.Median.df[,2:ncol(ExtractionDarkByMapQ.Median.df)]))

ggplot(extractionDarkByMapQ.Avg.melt[which(extractionDarkByMapQ.Avg.melt$value != "."),], aes(x=log10(as.numeric(as.character(value))), color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("log10(Average Depth)") + ggtitle("Average Depth of Dark-By-MapQ Extraction Regions")

#ggplot(extractionDarkByMapQ.Median.melt[which(extractionDarkByMapQ.Median.melt$value != "."),], aes(x=log10(as.numeric(as.character(value))), color=Var2)) + geom_density() + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("log10(Median Depth)") + ggtitle("Median Depth of Dark-By-MapQ Extraction Regions")
```

# All Dark Region Analyses

## Load Dark Region Data

```{r}
AllDarkByDepthFiles = Sys.glob(paste0(params$calculateBamStats, "/*.low_depth-merged.bed"))
AllDarkByMapQFiles = Sys.glob(paste0(params$calculateBamStats, "/*.low_mapq-merged.bed"))

CumulativeDarkByDepthBases = NULL
CumulativeDarkByMapQBases = NULL

AllDarkByDepth.Avg.dt = NULL
AllDarkByMapQ.Avg.dt = NULL

AllDarkByDepth.Median.dt = NULL
AllDarkByMapQ.Median.dt = NULL

for( file in AllDarkByDepthFiles ){
  temp = fread(file, sep = "\t", header = F)
  temp$rn = paste0(temp$V1,":",temp$V2,"-",temp$V3)
  setkey(temp, rn)
  
  if(!is.null(AllDarkByDepth.Avg.dt)){
    newColName = c(colnames(AllDarkByDepth.Avg.dt), unlist(strsplit(basename(file), "_"))[1])
    AllDarkByDepth.Avg.dt = merge(AllDarkByDepth.Avg.dt, temp[,c("rn", "V4")], by="rn", all=T)
    colnames(AllDarkByDepth.Avg.dt) = newColName
    
    AllDarkByDepth.Median.dt = merge(AllDarkByDepth.Median.dt, temp[,c("rn", "V5")], by="rn", all=T)
    colnames(AllDarkByDepth.Median.dt) = newColName
    
    CumulativeDarkByDepthBases = rbind(CumulativeDarkByDepthBases, c(unlist(strsplit(basename(file), "_"))[1], sum(temp$V3-temp$V2)))
  } else{
    AllDarkByDepth.Avg.dt = temp[,c("rn","V4")]
    colnames(AllDarkByDepth.Avg.dt) = c("rn", unlist(strsplit(basename(file), "_"))[1])
    
    AllDarkByDepth.Median.dt = temp[,c("rn","V5")]
    colnames(AllDarkByDepth.Median.dt) = c("rn", unlist(strsplit(basename(file), "_"))[1])
    
    CumulativeDarkByDepthBases = data.frame(Sample = c(unlist(strsplit(basename(file), "_"))[1]), CumBases = c(sum(temp$V3-temp$V2)))
  }
  
}

for( file in AllDarkByMapQFiles ){
  temp = fread(file, sep = "\t", header = F)
  temp$rn = paste0(temp$V1,":",temp$V2,"-",temp$V3)
  setkey(temp, rn)
  if(!is.null(AllDarkByMapQ.Avg.dt)){
    newColName = c(colnames(AllDarkByMapQ.Avg.dt), unlist(strsplit(basename(file), "_"))[1])
    AllDarkByMapQ.Avg.dt = merge(AllDarkByMapQ.Avg.dt, temp[,c("rn", "V4")], by="rn", all=T)
    colnames(AllDarkByMapQ.Avg.dt) = newColName
    
    AllDarkByMapQ.Median.dt = merge(AllDarkByMapQ.Median.dt, temp[,c("rn", "V5")], by="rn", all=T)
    colnames(AllDarkByMapQ.Median.dt) = newColName
    
    CumulativeDarkByMapQBases = rbind(CumulativeDarkByMapQBases, c(unlist(strsplit(basename(file), "_"))[1], sum(temp$V3-temp$V2)))
  } else{
    AllDarkByMapQ.Avg.dt = temp[,c("rn","V4")]
    colnames(AllDarkByMapQ.Avg.dt) = c("rn", unlist(strsplit(basename(file), "_"))[1])
    
    AllDarkByMapQ.Median.dt = temp[,c("rn","V5")]
    colnames(AllDarkByMapQ.Median.dt) = c("rn", unlist(strsplit(basename(file), "_"))[1])
    
    CumulativeDarkByMapQBases = data.frame(Sample = c(unlist(strsplit(basename(file), "_"))[1]), CumBases = c(sum(temp$V3-temp$V2)))
  }
  
}

AllDarkByMapQ.Avg.df = data.frame(AllDarkByMapQ.Avg.dt, row.names = "rn")
AllDarkByMapQ.Median.df = data.frame(AllDarkByMapQ.Median.dt, row.names = "rn")
AllDarkByDepth.Avg.df = data.frame(AllDarkByDepth.Avg.dt, row.names = "rn")
AllDarkByDepth.Median.df = data.frame(AllDarkByDepth.Median.dt, row.names = "rn")
```

## Number of Dark-by-MAPQ Regions vs. Dark-by-Depth Regions

```{r}
NumDark = data.frame(DarkByMapQ = sapply(1:ncol(AllDarkByMapQ.Avg.df), function(i) length(which(!is.na(as.numeric(as.character(AllDarkByMapQ.Avg.df[,i])))))), DarkByDepth = sapply(1:ncol(AllDarkByDepth.Avg.df), function(i) length(which(!is.na(as.numeric(as.character(AllDarkByDepth.Avg.df[,i])))))))
rownames(NumDark) = colnames(AllDarkByMapQ.Avg.df)

ggplot(NumDark, aes(x=log10(DarkByMapQ), y= log10(DarkByDepth))) + geom_point() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank())

```

## Dark-by-MAPQ

### Cumulative Histogram

```{r}
ggplot(CumulativeDarkByMapQBases, aes(x=log10(as.numeric(CumBases)))) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + ggtitle("Cumulative Number of Dark-By-MapQ Bases") + xlab("log10(Cumulative Number of Bases Covered by Dark-By-MapQ Regions)") + ylab("Number of Samples")
#hist(log10(as.numeric(CumulativeDarkByMapQBases$CumBases)), main="Cumulative Number of Dark-By-MapQ Bases", xlab = "log10(Cumulative Number of Bases Covered by Dark-By-MapQ Regions)", ylab="Number of Samples")
```


### Average Depth

```{r}
for(i in 1:ncol(AllDarkByMapQ.Avg.df)){
  print(ggplot(AllDarkByMapQ.Avg.df, aes(x=log10(AllDarkByMapQ.Avg.df[,i]))) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("log10(Dark-By-MapQ Average Depth)") + ylab("Number of Dark-By-MapQ Regions") + ggtitle(paste(colnames(AllDarkByMapQ.Avg.df)[i], "Average Dark-By-MapQ Depth"))) + xlim(c(0,5))
  #hist(log10(AllDarkByMapQ.Avg.df[,i]), main = paste(colnames(AllDarkByMapQ.Avg.df)[i], "Average Dark-By-MapQ Depth"), xlab="log10(Dark-By-MapQ Average Depth)")
}


mapqAvgDepth = melt(as.matrix(AllDarkByMapQ.Avg.df))

ggplot(mapqAvgDepth, aes(x=log10(value), color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("log10(Average Depth)") + ggtitle("Average Depth of Dark-By-MapQ per Extraction Region per Sample") + xlim(c(0,5))
```
### Median Depth

```{r eval=F}
for(i in 1:ncol(AllDarkByMapQ.Median.df)){
  hist(log10(AllDarkByMapQ.Median.df[,i]), main = paste(colnames(AllDarkByMapQ.Median.df)[i], "Median Dark-By-MapQ Depth"), xlab="log10(Dark-By-MapQ Median Depth)")
}

mapqMedDepth = melt(as.matrix(AllDarkByMapQ.Median.df))

ggplot(mapqMedDepth, aes(x=log10(value), color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("log10(Median Depth)") + ggtitle("Median Depth of Dark-By-MapQ per Extraction Region per Sample")
```





## Dark-by-Depth

### Cumulative Bases

```{r}
ggplot(CumulativeDarkByDepthBases, aes(x=log10(as.numeric(CumBases)))) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + ggtitle("Cumulative Number of Dark-By-Depth Bases") + xlab("log10(Cumulative Number of Bases Covered by Dark-By-Depth Regions)") + ylab("Number of Samples")
#hist(log10(as.numeric(CumulativeDarkByDepthBases$CumBases)), main="Cumulative Number of Dark-By-Depth Bases", xlab = "log10(Cumulative Number of Bases Covered by Dark-By-Depth)", ylab = "Number of Samples")
```

### Average Depth

```{r}
for(i in 1:ncol(AllDarkByDepth.Avg.df)){
  print(ggplot(AllDarkByDepth.Avg.df, aes(x=AllDarkByDepth.Avg.df[,i])) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Dark-By-Depth Average Depth") + ylab("Number of Dark-By-Depth Regions") + ggtitle(paste(colnames(AllDarkByDepth.Avg.df)[i], "Average Dark-By-Depth Depth")))
  #hist(AllDarkByDepth.Avg.df[,i], main = paste(colnames(AllDarkByDepth.Avg.df)[i], "Average Dark-By-Depth Depth"), xlab="log10(Dark-By-Depth Average Depth)")
}

depthMedDepth = melt(as.matrix(AllDarkByDepth.Avg.df))

ggplot(depthMedDepth, aes(x=value, color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Average Depth") + ggtitle("Average Depth of Dark-By-Depth per Extraction Region")
```


### Median Depth

```{r eval=F}
for(i in 1:ncol(AllDarkByDepth.Median.df)){
  hist(AllDarkByDepth.Median.df[,i], main = paste(colnames(AllDarkByDepth.Median.df)[i], "Median Dark-By-Depth Depth"), xlab="log10(Dark-By-Depth Median Depth)")
}

depthMedDepth = melt(as.matrix(AllDarkByDepth.Median.df))

ggplot(depthMedDepth, aes(x=value, color=Var2)) + geom_density() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Median Depth") + ggtitle("Median Depth of Dark-By-Depth per Extraction Region")
```


# Coverage of Genes of Interest

To get this data, we run samtools view on the Cram files to pull the gene regions. We then run bedtools genomecov to get the base by base coverage, which we then merge with the genome bed to get the gene name. That resultant bedgraph file is used in this code.

Relevant Code:

 * modules/01-RUN_DRF.nf : 183-208

```{r eval=T}
region = strsplit(rownames(AllDarkByDepth.Avg.df), ":|-")
AllDarkByDepth.Avg.df$chr = unlist(lapply(region, `[`, 1))
AllDarkByDepth.Avg.df$start = as.numeric(as.character(unlist(lapply(region, `[`, 2))))
AllDarkByDepth.Avg.df$end = as.numeric(as.character(unlist(lapply(region, `[`, 3))))

region = strsplit(rownames(AllDarkByMapQ.Avg.df), ":|-")
AllDarkByMapQ.Avg.df$chr = unlist(lapply(region, `[`, 1))
AllDarkByMapQ.Avg.df$start = as.numeric(as.character(unlist(lapply(region, `[`, 2))))
AllDarkByMapQ.Avg.df$end = as.numeric(as.character(unlist(lapply(region, `[`, 3))))
  
countBasesDarkByDepth = function(chr, start, end, sample){
  temp = AllDarkByDepth.Avg.df[which(AllDarkByDepth.Avg.df$chr == chr & start <= AllDarkByDepth.Avg.df$start & end >= AllDarkByDepth.Avg.df$end & !is.na(AllDarkByDepth.Avg.df[,sample])), c(sample, "start", "end"), drop=F]
  return(sum(temp$end-temp$start))
  
}

countBasesDarkByMapQ = function(chr, start, end, sample){
  temp = AllDarkByMapQ.Avg.df[which(AllDarkByMapQ.Avg.df$chr == chr & start <= AllDarkByMapQ.Avg.df$start & end >= AllDarkByMapQ.Avg.df$end & !is.na(AllDarkByMapQ.Avg.df[,sample])), c(sample, "start", "end"), drop=F]
  return(sum(temp$end-temp$start))
  
}

GoI = read.table(params$genesOfInterest, header=F, sep="\t")
Genes = unique(GoI[which(GoI$V4 !="."), "V4"])

assembledOnly = c("chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22", "chrX", "chrY", "chrM")



GenesOfInterestFiles = Sys.glob(paste0(params$runDRFDir, "/*cov.perBase.GOI.bed"))

for(gene in Genes){
  samples = NULL
  for(file in GenesOfInterestFiles){
    
    geneCov = fread(file, sep = "\t", header = F, stringsAsFactors = F)
    geneCov.sub = geneCov[which(geneCov$V8 == gene), c(2,4),drop=F]
    sample=unlist(strsplit(basename(file), "_"))[1]
    colnames(geneCov.sub) = c("rn", sample)
    
    if(!is.null(samples)){
      samples = merge(samples, geneCov.sub, all=T, by="rn")
    } else {
      samples = geneCov.sub
    }
    
    
    
  }
  
  samples.df = as.data.frame(samples)
  rownames(samples.df) = samples.df$rn
  samples.df = samples.df[-1]
  samples.melt = reshape2::melt(as.matrix(samples.df))
  
  chrom=unname(unlist(unique(GoI[which(GoI$V4 == gene), "V1"])))
  startPos=unname(unlist(unique(GoI[which(GoI$V4 == gene), "V2"])))
  endPos=unname(unlist(unique(GoI[which(GoI$V4 == gene), "V3"])))
  
  ##TODO: Add gene diagram and possibly make it hoverable
  mainplot = ggplot(samples.melt, aes(Var1, value)) + geom_smooth(se=F, span=.5,method = "loess", aes(color=Var2)) + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + ggtitle(gene) + xlab(unique(geneCov[which(geneCov$V8 == gene), "V1"])) + ylab("Coverage per Sample")
  
  write.table(samples.df, file=paste0("./", gene, "_AllSamples.txt"), sep="\t", quote = F)
  
  stats = data.frame(mean = apply(samples.df, 2, mean, na.rm=T), median = apply(samples.df, 2, median, na.rm=T), stdev = apply(samples.df, 2, sd, na.rm=T), numDarkByDepthBases = sapply(colnames(samples.df), function(s) countBasesDarkByDepth(chrom, startPos, endPos, gsub("-",".",s))), numDarkByMapQBases = 
sapply(colnames(samples.df), function(s) countBasesDarkByMapQ(chrom, startPos, endPos, gsub("-",".",s))))

  p1 = ggplot(stats, aes(x=mean)) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title = element_blank()) + xlab("Mean Coverage") + ylab("Number of Samples")
  
  p2 = ggplot(stats, aes(x=median)) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Median Coverage") + ylab("Number of Samples")
  
  p3 = ggplot(stats, aes(x=numDarkByDepthBases)) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank()) + xlab("Number of Bases Covered by Dark-By-Depth Regions") + ylab("Number of Samples")
  
  p4 = ggplot(stats, aes(x=numDarkByMapQBases)) + geom_histogram() + theme_bw() + theme(legend.position="none", panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), legend.title= element_blank())+ xlab("Number of Bases Covered by Dark-By-MapQ Regions") + ylab("Number of Samples")
  
  lay=rbind(c(1,1), c(2,3), c(4,5))
  
  jpeg(file = paste0("./", gene, ".Coverage.jpg"), height= 1000, width=1000)
  grid.arrange(mainplot, p1, p2, p3, p4, widths=c(2,2), heights=c(5,4,4), layout_matrix=lay)
  dev.off()
  
}

```


```{r results='asis', eval=T}

for (gene in Genes){
  cat(paste("##", gene))
  img=paste0("./", gene, ".Coverage.jpg")
  cat(paste("![img](./", img, sep='' ) , ")\n")

  }
cat("  \n")
```



